Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-02-13 Paper.Rmd
part done moving TL material to supp mat, in midst of editing the text describing the sims done for TL

We addressed Q3 using both models and data. An initial 
idea was presented in the Introduction that
the distribution (through time) of a spatially averaged quantity should be influenced by the 
nature of the dependence between the local quantities being averaged, including their
copula structure and tail dependence. For instance, suppose $E_i(t)$ are ecological
variables measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and suppose 
the distribution and temporal dynamics of the spatial average $\sum_i E_i(t)/N$ is of
interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
strongly right-tail dependent for all or most location pairs $i$ and $j$, then
exceptionally large values tend to occur at the same time in all or most locations.
This should tend to increase the skewness of the distribution through time of the
spatial average. Similarly, left-tail dependence between local variables
will tend to decreases skewness of the spatial mean. Skewness of the 
spatial mean corresponds to the "spikiness" of the spatial-mean time series. 
The spatial-mean time series may be the quantity of principal importance for 
cases such as pest or resource abundance or CO$_2$ or methane flux.

@@

Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we tested whether
the copula structure of dependence between population variables influences the
spatial version of Taylor's
law. Taylor's law is an extremely commonly observed and widely studied 
[@Taylor1961, @Taylor1988, @Cohen2013, @Xu2016] empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of these 
groups via a power law, $v_g = a \times m_g^b$, or equivalently via a log-log linear 
relationship, $\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So the means and variances in the Tayor's law formula are computed across 
space. Taylor's law has been 
verified in data from hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015,@Reuman2017].

@@

>>>>

We addressed Q3 using both models and data. An initial 
idea was presented in the Introduction that
the distribution (through time) of a spatially averaged quantity should be influenced by the 
nature of the dependence between the local quantities being averaged, including their
copula structure and tail dependence. For instance, suppose $E_i(t)$ are ecological
variables measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and suppose 
the distribution and temporal dynamics of the spatial average $\sum_i E_i(t)/N$ is of
interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
strongly right-tail dependent for all or most location pairs $i$ and $j$, then
exceptionally large values tend to occur at the same time in all or most locations.
This should tend to increase the skewness of the distribution through time of the
spatial average. Similarly, left-tail dependence between local variables
will tend to decreases skewness of the spatial mean. Skewness of the 
spatial mean corresponds to the "spikiness" of the spatial-mean time series or
its tendency to crash to low levels. 
The spatial-mean time series and its skewness may be quantities of 
principal importance for 
cases such as pest or resource abundance or CO$_2$ or methane flux.

@@

@@

<!-- 
This extincton risk material had been commented out in the sup mat:

Extinction risk results are not shown currently, it is introduced as a future idea in Discussion------

# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}

In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;

\begin{equation}\label{eq.extrisk_model}
P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
\end{equation}

$r$ is the liner growth rate and $K$ is the carrying capacity of the environment. $\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence. We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion). We simulated Eq. \ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;

- LC model (r=0) :

    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one. [see first subplot of Fig. \ref{fig_risk_local_disp} and Fig. \ref{fig_risk_global_disp}]

- Ricker model (r $\neq$ 0):

    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$. For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations. [see Fig. \ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]

    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]

With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig. \ref{fig_risk_numlocs_effect_local_disp} - Fig. \ref{fig_risk_numlocs_effect_global_disp}] for both model. -->

### 2019-02-13 SupportingInformation.Rmd
part done moving TL material to supp mat, in midst of editing the text describing the sims done for TL

# Calculating curvature for Taylor's law \label{TL_curvature}

<!--***DAN: Shya, I need a reminder how we did this, can you please tell me where the
code is, then I can write this section. Feel free to take a stab at it yourself, too,
though.
***Shya: What do you want to know? how did you calculate mean curvature and 
slope for made up data or you want to see code with real data?

@@

<!--***DAN: Dan has not edited this section yet at all-->
Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula. First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with co-variance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of time series, number of sampling locations and number of replica for that copula, respectively. Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal. With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D). We also tested statistically that whether linearity, homoskedasticity is maintained in each copula data set or not and plotted our results along with root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I). We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).

<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion------

# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}

In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;

\begin{equation}\label{eq.extrisk_model}
P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
\end{equation}

$r$ is the liner growth rate and $K$ is the carrying capacity of the environment. $\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence. We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion). We simulated Eq. \ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;

- LC model (r=0) :

    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one. [see first subplot of Fig. \ref{fig_risk_local_disp} and Fig. \ref{fig_risk_global_disp}]

- Ricker model (r $\neq$ 0):

    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$. For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations. [see Fig. \ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]

    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]

With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig. \ref{fig_risk_numlocs_effect_local_disp} - Fig. \ref{fig_risk_numlocs_effect_global_disp}] for both model. -->

>>>>

\noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of dependence between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is an extremely commonly observed and widely studied 
[@Taylor1961, @Taylor1988, @Cohen2013, @Xu2016] empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means and variances are computed across space. Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015,@Reuman2017].

We examined using simulations whether copula structure influences spatial Taylor's law as 
follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
with unit diagonal entries 
and off-diagonal entries $\rho=0.7$; $n=25$ was the number of locations at which populations
were sampled. Then $N \times 1000$ random vectors were generated from a multivariate normal 
distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
were organized into an $N \times n \times 1000$ array representing $1000$ simulations of 
populations measured in $n$ locations over $N$ time steps. 
The cumulative distribution function
(cdf) of the standard normal distribution was applied to these numbers, followed by the 
inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
These transformations made gamma distributed the marginal distributions of 
each population sampling location in each simulation. 
A skewed population marginal distribution is an 
important factor in Taylor's law [@Cohen2015,@Reuman2017]. We refer henceforth to the 
resulting $N \times n \times 1000$ array as our *N array* for this simulation exercise;
copula structure of dependences between populations in different sampling locations
was normal (N) for the simulations in this array.

We then considered an $n$-dimensional C copula with parameter chosen so that the 
Kendall correlations between each pair of the $n$ components were the same as those
of the N distribution constructed above.

the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula. First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with co-variance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of time series, number of sampling locations and number of replica for that copula, respectively. Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal. With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D). We also tested statistically that whether linearity, homoskedasticity is maintained in each copula data set or not and plotted our results along with root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I). We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).

<!--
Curvature calculations:
***DAN: Shya, I need a reminder how we did this, can you please tell me where the
code is, then I can write this section. Feel free to take a stab at it yourself, too,
though.
***Shya: What do you want to know? how did you calculate mean curvature and 
slope for made up data or you want to see code with real data?

@@
